---
title: "Practical Machine Learning - Course Project"
author: "Mandy"
date: "February 1, 2016"
output: html_document
---

##Project Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

The goal of this project is to create a machine-learning algorithm that can correctly identify the quality of barbell bicep curls by using data from belt, forearm, arm, and dumbbell monitors. There are five classifications of this exercise, one method is the correct form of the exercise while the other four are common mistakes: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).

<http://groupware.les.inf.puc-rio.br/har#ixzz3PO5pnm1R>

###Loading necessary packages
```{r, echo=FALSE}
library(caret)
library(randomForest)
library(knitr)
library(rpart)
```

###Getting and loading the data
```{r}
set.seed(12345)

training<-read.csv("pml-training.csv", header=TRUE)
testing<-read.csv("pml-testing.csv",header=TRUE)
```

####Partioning the training set into two  
60% for myTraining, 40% for myTesting
```{r}
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]
myTesting <- training[-inTrain, ]
dim(myTraining) 
dim(myTesting)
```

####Cleaning Data  
Remove NearZeroVariance variables
```{r}
nzv <- nearZeroVar(myTraining, saveMetrics=TRUE)
myTraining <- myTraining[,nzv$nzv==FALSE]

nzv<- nearZeroVar(myTesting,saveMetrics=TRUE)
myTesting <- myTesting[,nzv$nzv==FALSE]
```

Remove the first column of the myTraining data set
```{r}
myTraining <- myTraining[c(-1)]
```

Clean variables with more than 60% NA
```{r}
trainingV3 <- myTraining
for(i in 1:length(myTraining)) {
    if( sum( is.na( myTraining[, i] ) ) /nrow(myTraining) >= .7) {
        for(j in 1:length(trainingV3)) {
            if( length( grep(names(myTraining[i]), names(trainingV3)[j]) ) == 1)  {
                trainingV3 <- trainingV3[ , -j]
            }   
        } 
    }
}

# Set back to the original variable name
myTraining <- trainingV3
rm(trainingV3)
```

Transform the myTesting and testing data sets
```{r}
clean1 <- colnames(myTraining)
clean2 <- colnames(myTraining[, -58])  
myTesting <- myTesting[clean1]         
testing <- testing[clean2]           

dim(myTesting)
```

Coerce the data into the same type
```{r}
for (i in 1:length(testing) ) {
    for(j in 1:length(myTraining)) {
        if( length( grep(names(myTraining[i]), names(testing)[j]) ) == 1)  {
            class(testing[j]) <- class(myTraining[i])
        }      
    }      
}

# To get the same class between testing and myTraining
testing <- rbind(myTraining[2, -58] , testing)
testing <- testing[-1,]
```

###Prediction using Decision Tree
```{r}
set.seed(12345)
modFitA1 <- rpart(classe ~ ., data=myTraining, method="class")
```

Testing results using confusion matrix
```{r}
predictionsA1 <- predict(modFitA1, myTesting, type = "class")
cmtree <- confusionMatrix(predictionsA1, myTesting$classe)
cmtree
```

```{r}
plot(cmtree$table, col = cmtree$byClass, main = paste("Decision Tree Confusion Matrix: Accuracy =", round(cmtree$overall['Accuracy'], 4)))
```

###Prediction using Random Forests
```{r}
set.seed(12345)
modFitB1 <- randomForest(classe ~ ., data=myTraining)
```
```{r}
plot(modFitB1)
```

Testing results using confusion matrix
```{r}
predictionB1 <- predict(modFitB1, myTesting, type = "class")
cmrf <- confusionMatrix(predictionB1, myTesting$classe)
cmrf
```

```{r}
plot(cmrf$table, col = cmtree$byClass, main = paste("Random Forest Confusion Matrix: Accuracy =", round(cmrf$overall['Accuracy'], 4)))
```

####Conclude Result by using Test data
Random Forests gave an Accuracy in the myTesting dataset of 99.89%, which was more accurate that Decision Trees. 

```{r}
predictionB2 <- predict(modFitB1, testing, type = "class")
predictionB2
```


